Comparing Machine Learning algorithms for classifying cognitive states from fMRI data

### 2012-11-10 article/article.tex
boe

In these trials, subjects were presented with a stimulus on a computer screen. This stimulus was either a picture or a sentence. After a period of time, the stimulus would be replaced with a stimulus of the opposite type (i.e. picture with sentence and vice versa) after which the subject would indicate by pressing a button whether the sentence correctly described the picture. At the end of each trial, approximately 54 fMRI scans have be taken with an interval of 500ms.
The entire dataset consist of voxel data and metadata providing information about the dataset. Included in this information is not only the trial and scan each voxel belongs to, but also the type of the trial (picture first or sentence first). This makes the dataset suitable for classification using supervised learning.

In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject, 6 subjects. Voxels have been grouped into 25 regions of interest (see \textit{Feature selection}) that may be of use for analysis.

@@

Principal Component Analysis (PCA). As feature extraction method, we applied PCA. Ideally, the input vectors of the PCA should be the standardized Talairach coordinates, because we are conducting between-subject classification. As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices. The problem is that indices of different subjects may not map well onto each other, due to differences in brain size and shape. As a result, the accuracy of the PCA is reduced.

>>>>

In some of these trials (baseline trials), no stimulus was presented and participants were instructed to remain inert. In other trials, subjects were presented with a sequence of two stimuli on a computer screen. The first stimulus was either a picture or a sentence lasting 4 seconds, followed by a blank screen of 4 seconds. After the blank screen, the opposite type of stimulus was presented for 4 seconds, followed by a blank screen of 15 seconds, during which the subject would indicate by pressing a button whether the sentence correctly described the picture. During a trial, approximately 54 fMRI scans have be taken with an interval of 500ms. In total, 14 trials were baseline trials, while half of the remaining trials had a picture as first stimulus and the other half a sentence.  
The entire dataset consist of voxel data and metadata providing information about the dataset. Included in this information is not only the trial and scan each voxel belongs to, but also the type of the trial (picture first or sentence first). This makes the dataset suitable for classification using supervised learning.

In total, we have data available for approximately 4800 voxels per scan, 54 scans per trial, 40 trials per subject (excluding baseline trials) and 6 subjects. Voxels have been grouped into 25 regions of interest (see \textit{Feature selection}) that may be of use for analysis.

@@

Principal Component Analysis (PCA). As feature extraction method, we applied PCA. Ideally, the input vectors of the PCA should be the standardized Talairach coordinates, because we are conducting between-subject classification. As this coordinate system appeared to be inaccurate for this dataset, we were forced to use the voxel indices. The problem is that indices of different subjects may not map well onto each other, due to differences in brain size and shape. As a result, the accuracy of the PCA may be reduced.
